---
title: "An analysis of CO2 exchange rates compared with tree canopy height using CHM and eddy covariance data across three NEON study sites."
author: "McKenna Montminy"
date: "12/17/2020"
output: html_document
editor_options: 
  chunk_output_type: inline
---
First, load the libraries that we need for this study. 
```{r message=FALSE, warning=FALSE, results=FALSE}

# Load libraries
library(httr)
library(jsonlite)
library(dplyr)
library(devtools) 
library(downloader)
library(raster)
library(rgdal)
library(ggplot2)
library(dplyr)
library(sf)
library(tidyverse)
library(neonUtilities)
library(geoNEON)
library(ggplot2)
library(BiocManager)
library(rhdf5)
library(data.table)

```

Input eddy covariance data for ALL SITES [https://www.neonscience.org/resources/learning-hub/tutorials/eddy-data-intro]
```{r, setup/download data from NEON}

# Setup
options(stringsAsFactors=F)

# Download eddy covariance flux data
zipsByProduct(dpID="DP4.00200.001", package="basic", 
            site="SJER", 
            startdate="2019-10", enddate="2020-10")
zipsByProduct(dpID="DP4.00200.001", package="basic", 
            site="SOAP", 
            startdate="2019-10", enddate="2020-10")
zipsByProduct(dpID="DP4.00200.001", package="basic", 
            site="TEAK", 
            startdate="2019-10", enddate="2020-10")

flux <- stackEddy(filepath="filesToStack00200",
                 level="dp04") #extracting level 4 gives us flux data

```


## SJER
```{r}

# Eddy covariance data setup
SJER_term <- unlist(strsplit(names(flux$SJER), split=".", fixed=T))
flux$objDesc[which(flux$objDesc$Object %in% SJER_term),]

```

```{r, CO2 flux over a year}

## Eddy covariance: Entering annual time data
SJERtimeB <- as.POSIXct(flux$SJER$timeBgn, 
                    format="%Y-%m-%dT%H:%M:%S", 
                    tz="GMT")
flux$SJER <- cbind(SJERtimeB, flux$SJER)

# Plotted over year: Oct 2019 -- Oct 2020
plot(flux$SJER$data.fluxCo2.nsae.flux ~ SJERtimeB, 
     pch=".", xlab="Date (10/01/19 - 09/30/20)", ylab="CO2 flux",
     xlim=c(as.POSIXct("2019-10-01", tz="GMT"),
            as.POSIXct("2020-09-30", tz="GMT")),
    ylim=c(-20,20), xaxt="n") +
  axis.POSIXct(1, x=SJERtimeB, format="%Y-%m-%d")

```


```{r, CO2 flux over 1 day}

# Plotted over 1 day ***I chose APRIL 2020 because April is part of wet season with a lot of growth. 
plot(flux$SJER$data.fluxCo2.nsae.flux ~SJERtimeB, 
     pch=20, xlab="Date (04/01/20 - 04/03/20)", ylab="CO2 flux",
     xlim=c(as.POSIXct("2020-04-01", tz="GMT"),
            as.POSIXct("2020-04-03", tz="GMT")),
    ylim=c(-20,20), xaxt="n") +
  #abline(lm(flux$SJER$data.fluxCo2.nsae.flux ~ SJERtimeB), col = "blue") +
  axis.POSIXct(1, x=SJERtimeB, format="%Y-%m-%d %H:%M:%S")

```

### Overlay shapefile of eddy covariance tower footprint on NEON CHM data.
Read shapefile with the boundaries of a eddy covariance tower footprint.
```{r}

NEON_ECtower_fp <-sf::st_read("data/90percentfootprint/90percent_footprint.shp")

```

```{r}

# Filter to the SJER tower footprint
ECtower_SJER <- dplyr::filter(NEON_ECtower_fp, SiteID == "SJER")

# Check current CRS of the footprint shapefile
st_crs(ECtower_SJER)

```

```{r}

# Reproject the tower footprint crs from longlat to UTM 11N to match CHM footprint
# Define UTM 11N WGS84 proj4string
utm11nCRS <- "+proj=utm +zone=11 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
ECtower_SJER_utm <- st_transform(ECtower_SJER, utm11nCRS)

# Find limits of tower footprint to only download AOP tiles that are co-located with the tower footprint
SJER_bbox <- st_bbox(ECtower_SJER_utm)
SJER_bbox

# Round easting/northing to nearest 1km for matching AOP tiles that overlap tower footprint
SJER_easting <- c(256000, 256000, 256000, 257000, 257000, 257000, 258000, 258000, 258000)
SJER_northing <- c(4109000, 4110000, 4111000, 4109000, 4110000, 4111000, 4109000, 4110000, 4111000)

```

Request ecosystem structure data (CHM) from NEON API:
```{r}

# Download the canopy height tiles (1km grid) that correspond to tower footprint
SJERtmp <- byTileAOP(dpID = "DP3.30015.001", site = "SJER", year = c("2019"),
                 easting = SJER_easting, northing = SJER_northing)

# Load the canopy heights as raster files, and merge into 1 raster
SJER_chm_1 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SJER_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SJER_DP3_256000_4109000_CHM.tif")
SJER_chm_2 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SJER_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SJER_DP3_256000_4110000_CHM.tif")
SJER_chm_3 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SJER_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SJER_DP3_256000_4111000_CHM.tif")
SJER_chm_4 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SJER_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SJER_DP3_257000_4109000_CHM.tif")
SJER_chm_5 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SJER_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SJER_DP3_257000_4110000_CHM.tif")
SJER_chm_6 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SJER_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SJER_DP3_257000_4111000_CHM.tif")
SJER_chm_7 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SJER_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SJER_DP3_258000_4110000_CHM.tif")
SJER_chm_8 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SJER_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SJER_DP3_258000_4111000_CHM.tif")
SJER_chm_9 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SJER_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SJER_DP3_258000_4109000_CHM.tif")

# Merge individual 1km tiles
SJER_chm <- merge(SJER_chm_1, SJER_chm_2, SJER_chm_3, SJER_chm_4, 
                  SJER_chm_5, SJER_chm_6, SJER_chm_7, SJER_chm_8, SJER_chm_9)

# Check CHM crs
crs(SJER_chm)

```

### Plot tower footprint and CHM.
```{r}

# Make CHM raster a data frame
SJER_chm_df <- as.data.frame(SJER_chm, xy=TRUE)

# Plot the tower footprint shapefile onto the CHM
ggplot() +
  geom_raster(data = SJER_chm_df,
              aes(x = x, y = y, fill = layer)) + 
  geom_sf(data = ECtower_SJER_utm, color = "white", fill = NA) + 
  scale_fill_gradientn(name = "Canopy Height", colors = terrain.colors(10)) + 
  coord_sf(datum=utm11nCRS)

```

```{r, SJER summary: CHM mean & sd, and CO2 max/min}

# Extract CHM mean data values from shapefile footprint
SJER_fpCHM_mean <- raster::extract(x = SJER_chm, y = ECtower_SJER_utm, df = TRUE, fun=mean)
SJER_fpCHM_mean
## mean = 2.366575

# Extract CHM mstandard deviation data values from shapefile footprint
SJER_fpCHM_sd <- raster::extract(x = SJER_chm, y = ECtower_SJER_utm, df = TRUE, fun=sd)
SJER_fpCHM_sd
## sd = 4.393040

## Use ID #1; in ECtower_SJER under the notes column it states that ID #1 is the major site and #2 is secondary. So we'll use #1 to simplify. 

``` 

```{r, summary eddy covariance}

# Formatting flux list into a data frame to run stats on
SJER_CO2 <- as.data.frame(flux$SJER$data.fluxCo2.nsae.flux, as.is = TRUE)
SJER_CO2 <- type.convert(SJER_CO2)
colnames(SJER_CO2) <- "Co2_flux"
view(SJER_CO2)

# Filtering CO2 flux data to be within -100 to 100 (what is biologically possible)
SJER_CO2 <- dplyr::filter(SJER_CO2, Co2_flux <= 100)
SJER_CO2 <- dplyr::filter(SJER_CO2, Co2_flux > -100)
  
max(SJER_CO2, na.rm=TRUE)
# 95.10077
min(SJER_CO2, na.rm=TRUE)
# -96.48339

```



## SOAP

Input eddy covariance data [https://www.neonscience.org/resources/learning-hub/tutorials/eddy-data-intro]
```{r}

# Eddy covariance data setup
SOAP_term <- unlist(strsplit(names(flux$SOAP), split=".", fixed=T))
flux$objDesc[which(flux$objDesc$Object %in% SOAP_term),]

```

```{r, CO2 flux over (year)}

## Eddy covariance: Entering annual time data
SOAPtimeB <- as.POSIXct(flux$SOAP$timeBgn, 
                    format="%Y-%m-%dT%H:%M:%S", 
                    tz="GMT")
flux$SOAP <- cbind(SOAPtimeB, flux$SOAP)

# Plotted over year: Oct 2019 -- Oct 2020
plot(flux$SOAP$data.fluxCo2.nsae.flux ~ SOAPtimeB, 
     pch=".", xlab="Date (10/01/19 - 09/30/20)", ylab="CO2 flux",
     xlim=c(as.POSIXct("2019-10-01", tz="GMT"),
            as.POSIXct("2020-09-30", tz="GMT")),
    ylim=c(-20,20), xaxt="n")+
  axis.POSIXct(1, x=SOAPtimeB, format="%Y-%m-%d")

```

```{r, CO2 flux 1 day}

# Plotted over 1 day ***I chose APRIL 2020 because April is part of wet season with a lot of growth. 
plot(flux$SOAP$data.fluxCo2.nsae.flux ~SOAPtimeB, 
     pch=20, xlab="Date (04/01/20 - 04/03/20)", ylab="CO2 flux",
     xlim=c(as.POSIXct("2020-04-01", tz="GMT"),
            as.POSIXct("2020-03-03", tz="GMT")),
    ylim=c(-20,20), xaxt="n") +
  #abline(lm(flux$SOAP$data.fluxCo2.nsae.flux ~ SOAPtimeB), col = "blue") +
  axis.POSIXct(1, x=SOAPtimeB, format="%Y-%m-%d %H:%M:%S")

```

### Overlay shapefile of eddy covariance tower footprint on NEON CHM data.
Read shapefile with the boundaries of a eddy covariance tower footprint
```{r, SOAP tower footprint}

# Filter to the SOAP tower footprint
ECtower_SOAP <- dplyr::filter(NEON_ECtower_fp, SiteID == "SOAP")

# Check current CRS of the footprint shapefile
st_crs(ECtower_SOAP) 

```

```{r}
# Reproject the crs from longlat to UTM 11N
# Define UTM 11N WGS84 proj4string
utm11nCRS <- "+proj=utm +zone=11 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
ECtower_SOAP_utm <- st_transform(ECtower_SOAP, utm11nCRS)

# Find limits of tower footprint to only download AOP that is co-located with the footprint
SOAP_bbox <- st_bbox(ECtower_SOAP_utm)
SOAP_bbox

# Round easting/northing to nearest 1km for matching AOP tiles that overlap footprint
SOAP_easting <- c(297000, 297000, 298000, 298000, 299000, 299000)
SOAP_northing <- c(4100000, 4101000, 4100000, 4101000, 4100000, 4101000)

```

Request ecosystem structure data from NEON API:
```{r}
# Download the canopy height tiles (1km grid) that correspond to tower footprint
# This will download to a folder called DP3.30015.001 where your project directory is
SOAPtmp <- byTileAOP(dpID = "DP3.30015.001", site = "SOAP", year = c("2019", "2020"),
                 easting = SOAP_easting, northing = SOAP_northing)

# Load the canopy heights as raster files, and merge into 1 raster
SOAP_chm_1 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SOAP_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SOAP_DP3_297000_4100000_CHM.tif")
SOAP_chm_2 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SOAP_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SOAP_DP3_297000_4101000_CHM.tif")
SOAP_chm_3 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SOAP_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SOAP_DP3_298000_4100000_CHM.tif")
SOAP_chm_4 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SOAP_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SOAP_DP3_298000_4101000_CHM.tif")
SOAP_chm_5 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SOAP_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SOAP_DP3_299000_4100000_CHM.tif")
SOAP_chm_6 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SOAP_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SOAP_DP3_299000_4101000_CHM.tif")
   
SOAP_chm <- merge(SOAP_chm_1, SOAP_chm_2, SOAP_chm_3, SOAP_chm_4, SOAP_chm_5, SOAP_chm_6)

# Check CHM crs
crs(SOAP_chm)
```

### Plot tower footprint and CHM.
```{r}
# Make raster a data frame
SOAP_chm_df <- as.data.frame(SOAP_chm, xy=TRUE)

# Plot the tower footprint shapefile onto the CHM
ggplot() +
  geom_raster(data = SOAP_chm_df,
              aes(x = x, y = y, fill = layer)) + 
  geom_sf(data = ECtower_SOAP_utm, color = "white", fill = NA) + 
  scale_fill_gradientn(name = "Canopy Height", colors = terrain.colors(10)) + 
  coord_sf(datum=utm11nCRS)

```

```{r, SOAP CHM summary: mean & sd, echo = false}

# Extract CHM data values from the footprint shapefile
SOAP_fpCHM_mean <- raster::extract(x = SOAP_chm, y = ECtower_SOAP_utm, df = TRUE, fun=mean)
SOAP_fpCHM_mean
## mean = 5.561121

SOAP_fpCHM_sd <- raster::extract(x = SOAP_chm, y = ECtower_SOAP_utm, df = TRUE, fun=sd)
SOAP_fpCHM_sd
## sd = 8.232840

```

```{r, summary eddy covariance}

#Formatting flux list into a data frame to run stats on
SOAP_CO2 <- as.data.frame(flux$SOAP$data.fluxCo2.nsae.flux, as.is = TRUE)
SOAP_CO2 <- type.convert(SOAP_CO2)
colnames(SOAP_CO2) <- "Co2_flux"
view(SOAP_CO2)

# Filtering CO2 flux data to be within -100 to 100 (what is biologically possible)
SOAP_CO2 <- filter(SOAP_CO2, Co2_flux <= 100)
SOAP_CO2 <- filter(SOAP_CO2, Co2_flux > -100)
  
max(SOAP_CO2, na.rm=TRUE)
# 89.03265
min(SOAP_CO2, na.rm=TRUE)
# -96.4564

```



## TEAK

### Input eddy covariance data [https://www.neonscience.org/resources/learning-hub/tutorials/eddy-data-intro]
```{r}

# Eddy covariance data setup
TEAK_term <- unlist(strsplit(names(flux$TEAK), split=".", fixed=T))
flux$objDesc[which(flux$objDesc$Object %in% TEAK_term),]

```

```{r, CO2 flux over (time period? months?)}

## Eddy covariance: Entering annual time data
TEAKtimeB <- as.POSIXct(flux$TEAK$timeBgn, 
                    format="%Y-%m-%dT%H:%M:%S", 
                    tz="GMT")
flux$TEAK <- cbind(TEAKtimeB, flux$TEAK)

# Plotted over year: Oct 2019 -- Oct 2020
plot(flux$TEAK$data.fluxCo2.nsae.flux ~ TEAKtimeB, 
     pch=".", xlab="Date (10/01/19 - 09/30/20)", ylab="CO2 flux",
     xlim=c(as.POSIXct("2019-10-01", tz="GMT"),
            as.POSIXct("2020-09-30", tz="GMT")),
    ylim=c(-20,20), xaxt="n") +
  axis.POSIXct(1, x=TEAKtimeB, format="%Y-%m-%d")

```

```{r, CO2 flux 1 day}

# Plotted over 1 day ***I chose APRIL 2020 because April is part of wet season with a lot of growth. 
plot(flux$TEAK$data.fluxCo2.nsae.flux ~ TEAKtimeB, 
     pch=20, xlab="Date (04/01/2020 - 04/03/2020)", ylab="CO2 flux",
     xlim=c(as.POSIXct("2020-04-01", tz="GMT"),
            as.POSIXct("2020-04-03", tz="GMT")),
    ylim=c(-20,20), xaxt="n") +
  #abline(lm(flux$TEAK$data.fluxCo2.nsae.flux ~ TEAKtimeB), col = "blue") +
  axis.POSIXct(1, x=TEAKtimeB, format="%Y-%m-%d %H:%M:%S")

```

### Overlay shapefile of eddy covariance tower footprint on NEON CHM data.

```{r}

# Filter to the SJER tower footprint
ECtower_TEAK <- dplyr::filter(NEON_ECtower_fp, SiteID == "TEAK")

# Check current CRS of the footprint shapefile
st_crs(ECtower_TEAK) 

```

```{r}

# Reproject the crs from longlat to UTM 11N
# Define UTM 11N WGS84 proj4string
utm11nCRS <- "+proj=utm +zone=11 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
ECtower_TEAK_utm <- st_transform(ECtower_TEAK, utm11nCRS)

# Find limits of tower footprint to only download AOP that is co-located with the footprint
TEAK_bbox <- st_bbox(ECtower_TEAK_utm)
TEAK_bbox

# Actual footprint tiles
TEAK_easting <- c(320000, 320000, 320000, 321000, 321000, 321000, 322000, 322000, 322000)
TEAK_northing <- c(4096000, 4097000, 4098000, 4096000, 4097000, 4098000, 4096000, 4097000, 4098000)

```

Request ecosystem structure data from NEON API:
```{r}
# Download the canopy height tiles (1km grid) that correspond to tower footprint
# This will download to a folder called DP3.30015.001 where your project directory is
TEAKtmp <- byTileAOP(dpID = "DP3.30015.001", site = "TEAK", year = c("2019", "2020"),
                 easting = TEAK_easting, northing = TEAK_northing)

# Load the canopy heights as raster files, and merge into 1 raster
TEAK_chm_1 <- raster("DP3.30015.001/2019/FullSite/D17/2019_TEAK_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_TEAK_DP3_320000_4096000_CHM.tif")
TEAK_chm_2 <- raster("DP3.30015.001/2019/FullSite/D17/2019_TEAK_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_TEAK_DP3_320000_4097000_CHM.tif")
TEAK_chm_3 <- raster("DP3.30015.001/2019/FullSite/D17/2019_TEAK_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_TEAK_DP3_320000_4098000_CHM.tif")
TEAK_chm_4 <- raster("DP3.30015.001/2019/FullSite/D17/2019_TEAK_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_TEAK_DP3_321000_4096000_CHM.tif")
TEAK_chm_5 <- raster("DP3.30015.001/2019/FullSite/D17/2019_TEAK_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_TEAK_DP3_321000_4097000_CHM.tif")
TEAK_chm_6 <- raster("DP3.30015.001/2019/FullSite/D17/2019_TEAK_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_TEAK_DP3_321000_4098000_CHM.tif")
TEAK_chm_7 <- raster("DP3.30015.001/2019/FullSite/D17/2019_TEAK_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_TEAK_DP3_322000_4096000_CHM.tif")
TEAK_chm_8 <- raster("DP3.30015.001/2019/FullSite/D17/2019_TEAK_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_TEAK_DP3_322000_4097000_CHM.tif")
TEAK_chm_9 <- raster("DP3.30015.001/2019/FullSite/D17/2019_TEAK_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_TEAK_DP3_322000_4098000_CHM.tif")

TEAK_chm <- merge(TEAK_chm_1, TEAK_chm_2, TEAK_chm_3, TEAK_chm_4, 
                  TEAK_chm_5, TEAK_chm_6, TEAK_chm_7, TEAK_chm_8, TEAK_chm_9)

# Check CHM crs
crs(TEAK_chm)

```

### Plot tower footprint and CHM.
```{r}

# Make raster a data frame
TEAK_chm_df <- as.data.frame(TEAK_chm, xy=TRUE)

# Plot the tower footprint shapefile onto the CHM
ggplot() +
  geom_raster(data = TEAK_chm_df,
              aes(x = x, y = y, fill = layer)) + 
  geom_sf(data = ECtower_TEAK_utm, color = "white", fill = NA) + 
  scale_fill_gradientn(name = "Canopy Height", colors = terrain.colors(10)) + 
  coord_sf(datum=utm11nCRS)

```

```{r, TEAK CHM summary: mean & sd}

# Get CHM data values (mean & sd) inside EC footprint shapefile
TEAK_fpCHM_mean <- raster::extract(x = TEAK_chm, y = ECtower_TEAK_utm, df=TRUE, fun=mean)
TEAK_fpCHM_mean
## mean = 10.111631

TEAK_fpCHM_sd <- raster::extract(x = TEAK_chm, y = ECtower_TEAK_utm, df=TRUE, fun=sd)
TEAK_fpCHM_sd
## sd = 10.97886

```  

```{r, summary eddy covariance}

# Formatting flux list into a data frame to run stats on
TEAK_CO2 <- as.data.frame(flux$TEAK$data.fluxCo2.nsae.flux, as.is = TRUE)
TEAK_CO2 <- type.convert(TEAK_CO2)
colnames(TEAK_CO2) <- "Co2_flux"
view(TEAK_CO2)

# Filtering CO2 flux data to be within -100 to 100 (what is biologically possible)
TEAK_CO2 <- filter(TEAK_CO2, Co2_flux < "100")
TEAK_CO2 <- filter(TEAK_CO2, Co2_flux > -100)
  
max(TEAK_CO2, na.rm=TRUE)
# 10.96734
min(TEAK_CO2, na.rm=TRUE)
# -62.20244

```