---
title: "SJER, SOAP, & TEAK CO2 flux and CHM"
author: "McKenna Montminy"
date: "12/10/2020"
output: html_document
editor_options: 
  chunk_output_type: inline
---
First, load the libraries that we need for this study. 
```{r message=FALSE, warning=FALSE, results=FALSE}

# Load libraries
library(httr)
library(jsonlite)
library(dplyr)
library(devtools) 
library(downloader)
library(raster)
library(rgdal)
library(ggplot2)
library(dplyr)
library(sf)
library(tidyverse)
library(neonUtilities)
library(geoNEON)
library(ggplot2)
library(BiocManager)
library(rhdf5)

```

Input eddy covariance data for ALL SITES [https://www.neonscience.org/resources/learning-hub/tutorials/eddy-data-intro]
```{r, setup/download data from NEON}

# Setup
options(stringsAsFactors=F)

# Download data
#zipsByProduct(dpID="DP4.00200.001", package="basic", 
            #  site="SJER", 
            #  startdate="2019-10", enddate="2020-10")
#zipsByProduct(dpID="DP4.00200.001", package="basic", 
            #  site="SOAP", 
            #  startdate="2019-10", enddate="2020-10")
#zipsByProduct(dpID="DP4.00200.001", package="basic", 
            #  site="TEAK", 
            #  startdate="2019-10", enddate="2020-10")

flux <- stackEddy(filepath="data/filesToStack00200/",
                 level="dp04") 
#extracting level 4 gives us flux data
names(flux)

```


## SJER
```{r}

SJER_term <- unlist(strsplit(names(flux$SJER), split=".", fixed=T))
flux$objDesc[which(flux$objDesc$Object %in% SJER_term),]
head(SJER_term)
flux$variables

```

```{r, CO2 flux over (time period? months?)}

## Eddy [contd.]: Entering time data
SJERtimeB <- as.POSIXct(flux$SJER$timeBgn, 
                    format="%Y-%m-%dT%H:%M:%S", 
                    tz="GMT")
flux$SJER <- cbind(SJERtimeB, flux$SJER)

# Plotted over year: Oct 2019 -- Oct 2020
plot(flux$SJER$data.fluxCo2.nsae.flux ~ SJERtimeB, 
     pch=".", xlab="Date (10/01/20 - 09/30/20)", ylab="CO2 flux",
     xlim=c(as.POSIXct("2019-10-01", tz="GMT"),
            as.POSIXct("2020-09-30", tz="GMT")),
    ylim=c(-2000,2000), xaxt="n") +
  axis.POSIXct(1, x=SJERtimeB, format="%Y-%m-%d")

```


```{r, CO2 flux 1 day}

# Plotted over 1 day
plot(flux$SJER$data.fluxCo2.nsae.flux ~SJERtimeB, 
     pch=20, xlab="Date (01/01/20 - 01/03/20)", ylab="CO2 flux",
     xlim=c(as.POSIXct("2020-01-01", tz="GMT"),
            as.POSIXct("2020-01-03", tz="GMT")),
    ylim=c(-20,20), xaxt="n") +
  abline(lm(flux$SJER$data.fluxCo2.nsae.flux ~ SJERtimeB), col = "blue") +
  axis.POSIXct(1, x=SJERtimeB, format="%Y-%m-%d %H:%M:%S")

```

### Overlay shapefile of eddy covariance tower footprint on NEON CHM data.
Read shapefile with the boundaries of a eddy covariance tower footprint
```{r}

NEON_ECtower_fp <-sf::st_read("data/90percentfootprint/90percent_footprint.shp")

```

```{r}

# Filter to the SJER tower footprint
ECtower_SJER <- dplyr::filter(NEON_ECtower_fp, SiteID == "SJER")

# Check current CRS of the footprint shapefile
st_crs(ECtower_SJER)

```

```{r}

# Reproject the crs from longlat to UTM 11N
# Define UTM 11N WGS84 proj4string
utm11nCRS <- "+proj=utm +zone=11 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
ECtower_SJER_utm <- st_transform(ECtower_SJER, utm11nCRS)

# Find limits of tower footprint to only download AOP that is co-located with the footprint
SJER_bbox <- st_bbox(ECtower_SJER_utm)
SJER_bbox

# Round easting/northing to nearest 1km for matching AOP tiles that overlap footprint
SJER_easting <- c(256000, 256000, 256000, 257000, 257000, 257000, 258000, 258000, 258000)
SJER_northing <- c(4109000, 4110000, 4111000, 4109000, 4110000, 4111000, 4109000, 4110000, 4111000)

```

Request ecosystem structure data from NEON API:
```{r}

# Download the canopy height tiles (1km grid) that correspond to tower footprint
SJERtmp <- byTileAOP(dpID = "DP3.30015.001", site = "SJER", year = c("2019"),
                 easting = SJER_easting, northing = SJER_northing)

# Load the canopy heights as raster files, and merge into 1 raster
SJER_chm_1 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SJER_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SJER_DP3_256000_4109000_CHM.tif")
SJER_chm_2 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SJER_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SJER_DP3_256000_4110000_CHM.tif")
SJER_chm_3 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SJER_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SJER_DP3_256000_4111000_CHM.tif")
SJER_chm_4 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SJER_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SJER_DP3_257000_4109000_CHM.tif")
SJER_chm_5 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SJER_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SJER_DP3_257000_4110000_CHM.tif")
SJER_chm_6 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SJER_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SJER_DP3_257000_4111000_CHM.tif")
SJER_chm_7 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SJER_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SJER_DP3_258000_4110000_CHM.tif")
SJER_chm_8 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SJER_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SJER_DP3_258000_4111000_CHM.tif")
SJER_chm_9 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SJER_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SJER_DP3_258000_4109000_CHM.tif")

# Merge individual 1km tiles
SJER_chm <- merge(SJER_chm_1, SJER_chm_2, SJER_chm_3, SJER_chm_4, 
                  SJER_chm_5, SJER_chm_6, SJER_chm_7, SJER_chm_8, SJER_chm_9)

# Check CHM crs
crs(SJER_chm)

```
### Plot tower footprint and CHM.
```{r}

# Make raster a data frame
SJER_chm_df <- as.data.frame(SJER_chm, xy=TRUE)

# Plot the tower footprint shapefile onto the CHM
ggplot() +
  geom_raster(data = SJER_chm_df,
              aes(x = x, y = y, fill = layer)) + 
  geom_sf(data = ECtower_SJER_utm, color = "white", fill = NA) + 
  scale_fill_gradientn(name = "Canopy Height", colors = terrain.colors(10)) + 
  coord_sf(datum=utm11nCRS)

```

```{r, SJER summary: CHM mean & sd, and CO2 max/min}

# Extract CHM mean data values from shapefile footprint
SJER_fpCHM_mean <- raster::extract(x = SJER_chm, y = ECtower_SJER_utm, df = TRUE, fun=mean)
SJER_fpCHM_mean
## mean = 2.366575

# Extract CHM mstandard deviation data values from shapefile footprint
SJER_fpCHM_sd <- raster::extract(x = SJER_chm, y = ECtower_SJER_utm, df = TRUE, fun=sd)
SJER_fpCHM_sd
## sd = 4.393040

## Use ID #1; in ECtower_SJER under the notes column it states that ID #1 is the major site and #2 is secondary. So we'll use #1 to simplify. 

``` 

```{r, summary eddy covariance}

max(flux$SJER$data.fluxCo2.nsae.flux, na.rm=TRUE)
# 3306963
min(flux$SJER$data.fluxCo2.nsae.flux, na.rm=TRUE)
# -36934.04
mean(flux$SJER$data.fluxCo2.nsae.flux, na.rm=TRUE)
# 389.3742

```



## SOAP

Input eddy covariance data [https://www.neonscience.org/resources/learning-hub/tutorials/eddy-data-intro]
```{r}

SOAP_term <- unlist(strsplit(names(flux$SOAP), split=".", fixed=T))
flux$objDesc[which(flux$objDesc$Object %in% SOAP_term),]
head(SOAP_term)
flux$variables

```

```{r, CO2 flux over (year)}

SOAPtimeB <- as.POSIXct(flux$SOAP$timeBgn, 
                    format="%Y-%m-%dT%H:%M:%S", 
                    tz="GMT")
flux$SOAP <- cbind(SOAPtimeB, flux$SOAP)

# Plotted over year: Oct 2019 -- Oct 2020
plot(flux$SOAP$data.fluxCo2.nsae.flux ~ SOAPtimeB, 
     pch=".", xlab="Date (10/01/20 - 09/30/20)", ylab="CO2 flux",
     xlim=c(as.POSIXct("2019-10-01", tz="GMT"),
            as.POSIXct("2020-09-30", tz="GMT")),
    ylim=c(-2000,2000), xaxt="n")+
  axis.POSIXct(1, x=SOAPtimeB, format="%Y-%m-%d")

```

```{r, CO2 flux 1 day}

# Plotted over 1 day ***Changed to Jan 2020 because little data in Dec. 2019 & no data in Nov. 2019
plot(flux$SOAP$data.fluxCo2.nsae.flux ~SOAPtimeB, 
     pch=20, xlab="Date", ylab="CO2 flux",
     xlim=c(as.POSIXct("2020-01-01", tz="GMT"),
            as.POSIXct("2020-01-03", tz="GMT")),
    ylim=c(-20,20), xaxt="n") +
  abline(lm(flux$SOAP$data.fluxCo2.nsae.flux ~ SOAPtimeB), col = "blue") +
  axis.POSIXct(1, x=SOAPtimeB, format="%Y-%m-%d %H:%M:%S")

```

### Overlay shapefile of eddy covariance tower footprint on NEON CHM data.
Read shapefile with the boundaries of a eddy covariance tower footprint
```{r, SOAP tower footprint}

# Filter to the SOAP tower footprint
ECtower_SOAP <- dplyr::filter(NEON_ECtower_fp, SiteID == "SOAP")

# Check current CRS of the footprint shapefile
st_crs(ECtower_SOAP) 

```

```{r}
# Reproject the crs from longlat to UTM 11N
# Define UTM 11N WGS84 proj4string
utm11nCRS <- "+proj=utm +zone=11 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
ECtower_SOAP_utm <- st_transform(ECtower_SOAP, utm11nCRS)

# Find limits of tower footprint to only download AOP that is co-located with the footprint
SOAP_bbox <- st_bbox(ECtower_SOAP_utm)
SOAP_bbox

# Round easting/northing to nearest 1km for matching AOP tiles that overlap footprint
SOAP_easting <- c(297000, 297000, 298000, 298000, 299000, 299000)
SOAP_northing <- c(4100000, 4101000, 4100000, 4101000, 4100000, 4101000)

```

Request ecosystem structure data from NEON API:
```{r}
# Download the canopy height tiles (1km grid) that correspond to tower footprint
# This will download to a folder called DP3.30015.001 where your project directory is
SOAPtmp <- byTileAOP(dpID = "DP3.30015.001", site = "SOAP", year = c("2019", "2020"),
                 easting = SOAP_easting, northing = SOAP_northing)

# Load the canopy heights as raster files, and merge into 1 raster
SOAP_chm_1 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SOAP_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SOAP_DP3_297000_4100000_CHM.tif")
SOAP_chm_2 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SOAP_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SOAP_DP3_297000_4101000_CHM.tif")
SOAP_chm_3 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SOAP_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SOAP_DP3_298000_4100000_CHM.tif")
SOAP_chm_4 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SOAP_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SOAP_DP3_298000_4101000_CHM.tif")
SOAP_chm_5 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SOAP_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SOAP_DP3_299000_4100000_CHM.tif")
SOAP_chm_6 <- raster("DP3.30015.001/2019/FullSite/D17/2019_SOAP_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_SOAP_DP3_299000_4101000_CHM.tif")

SOAP_chm <- merge(SOAP_chm_1, SOAP_chm_2, SOAP_chm_3, SOAP_chm_4, SOAP_chm_5, SOAP_chm_6)

# Check CHM crs
crs(SOAP_chm)
```

### Plot tower footprint and CHM.
```{r}
# Make raster a data frame
SOAP_chm_df <- as.data.frame(SOAP_chm, xy=TRUE)

# Plot the tower footprint shapefile onto the CHM
ggplot() +
  geom_raster(data = SOAP_chm_df,
              aes(x = x, y = y, fill = layer)) + 
  geom_sf(data = ECtower_SOAP_utm, color = "white", fill = NA) + 
  scale_fill_gradientn(name = "Canopy Height", colors = terrain.colors(10)) + 
  coord_sf(datum=utm11nCRS)

```

```{r, SOAP CHM summary: mean & sd, echo = false}

# Extract CHM data values from the footprint shapefile
SOAP_fpCHM_mean <- raster::extract(x = SOAP_chm, y = ECtower_SOAP_utm, df = TRUE, fun=mean)
SOAP_fpCHM_mean
## mean = 5.561121

SOAP_fpCHM_sd <- raster::extract(x = SOAP_chm, y = ECtower_SOAP_utm, df = TRUE, fun=sd)
SOAP_fpCHM_sd
## sd = 8.232840

```

```{r, summary eddy covariance}

max(flux$SOAP$data.fluxCo2.nsae.flux, na.rm=TRUE)
# 256.364
min(flux$SOAP$data.fluxCo2.nsae.flux, na.rm=TRUE)
# -549.747
mean(flux$SOAP$data.fluxCo2.nsae.flux, na.rm=TRUE)
# -0.6100276

```




## TEAK

### Input eddy covariance data [https://www.neonscience.org/resources/learning-hub/tutorials/eddy-data-intro]
```{r}

TEAK_term <- unlist(strsplit(names(flux$TEAK), split=".", fixed=T))
flux$objDesc[which(flux$objDesc$Object %in% TEAK_term),]
head(TEAK_term)
flux$variables

```

```{r, CO2 flux over (time period? months?)}

TEAKtimeB <- as.POSIXct(flux$TEAK$timeBgn, 
                    format="%Y-%m-%dT%H:%M:%S", 
                    tz="GMT")
flux$TEAK <- cbind(TEAKtimeB, flux$TEAK)

# Plotted over year: Oct 2019 -- Oct 2020
plot(flux$TEAK$data.fluxCo2.nsae.flux ~ TEAKtimeB, 
     pch=".", xlab="Date (10/01/20 - 09/30/20)", ylab="CO2 flux",
     xlim=c(as.POSIXct("2019-10-01", tz="GMT"),
            as.POSIXct("2020-09-30", tz="GMT")),
    ylim=c(-2000,2000), xaxt="n") +
  axis.POSIXct(1, x=TEAKtimeB, format="%Y-%m-%d")

```

```{r, CO2 flux 1 day}
# Plotted over 1 day
## Changed to 12-02 to 12-04 because no data 12-01
plot(flux$TEAK$data.fluxCo2.nsae.flux ~ TEAKtimeB, 
     pch=20, xlab="Date", ylab="CO2 flux",
     xlim=c(as.POSIXct("2019-12-02", tz="GMT"),
            as.POSIXct("2019-12-04", tz="GMT")),
    ylim=c(-20,20), xaxt="n") +
  abline(lm(flux$TEAK$data.fluxCo2.nsae.flux ~ TEAKtimeB), col = "blue") +
  axis.POSIXct(1, x=TEAKtimeB, format="%Y-%m-%d %H:%M:%S")
```

### Overlay shapefile of eddy covariance tower footprint on NEON CHM data.

```{r}

# Filter to the SJER tower footprint
ECtower_TEAK <- dplyr::filter(NEON_ECtower_fp, SiteID == "TEAK")

# Check current CRS of the footprint shapefile
st_crs(ECtower_TEAK) 

```

```{r}

# Reproject the crs from longlat to UTM 11N
# Define UTM 11N WGS84 proj4string
utm11nCRS <- "+proj=utm +zone=11 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
ECtower_TEAK_utm <- st_transform(ECtower_TEAK, utm11nCRS)

# Find limits of tower footprint to only download AOP that is co-located with the footprint
TEAK_bbox <- st_bbox(ECtower_TEAK_utm)
TEAK_bbox

# Actual footprint tiles
TEAK_easting <- c(320000, 320000, 320000, 321000, 321000, 321000, 322000, 322000, 322000)
TEAK_northing <- c(4096000, 4097000, 4098000, 4096000, 4097000, 4098000, 4096000, 4097000, 4098000)

```

Request ecosystem structure data from NEON API:
```{r}
# Download the canopy height tiles (1km grid) that correspond to tower footprint
# This will download to a folder called DP3.30015.001 where your project directory is
TEAKtmp <- byTileAOP(dpID = "DP3.30015.001", site = "TEAK", year = c("2019", "2020"),
                 easting = TEAK_easting, northing = TEAK_northing)

# Load the canopy heights as raster files, and merge into 1 raster
TEAK_chm_1 <- raster("DP3.30015.001/2019/FullSite/D17/2019_TEAK_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_TEAK_DP3_320000_4096000_CHM.tif")
TEAK_chm_2 <- raster("DP3.30015.001/2019/FullSite/D17/2019_TEAK_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_TEAK_DP3_320000_4097000_CHM.tif")
TEAK_chm_3 <- raster("DP3.30015.001/2019/FullSite/D17/2019_TEAK_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_TEAK_DP3_320000_4098000_CHM.tif")
TEAK_chm_4 <- raster("DP3.30015.001/2019/FullSite/D17/2019_TEAK_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_TEAK_DP3_321000_4096000_CHM.tif")
TEAK_chm_5 <- raster("DP3.30015.001/2019/FullSite/D17/2019_TEAK_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_TEAK_DP3_321000_4097000_CHM.tif")
TEAK_chm_6 <- raster("DP3.30015.001/2019/FullSite/D17/2019_TEAK_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_TEAK_DP3_321000_4098000_CHM.tif")
TEAK_chm_7 <- raster("DP3.30015.001/2019/FullSite/D17/2019_TEAK_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_TEAK_DP3_322000_4096000_CHM.tif")
TEAK_chm_8 <- raster("DP3.30015.001/2019/FullSite/D17/2019_TEAK_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_TEAK_DP3_322000_4097000_CHM.tif")
TEAK_chm_9 <- raster("DP3.30015.001/2019/FullSite/D17/2019_TEAK_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D17_TEAK_DP3_322000_4098000_CHM.tif")

TEAK_chm <- merge(TEAK_chm_1, TEAK_chm_2, TEAK_chm_3, TEAK_chm_4, 
                  TEAK_chm_5, TEAK_chm_6, TEAK_chm_7, TEAK_chm_8, TEAK_chm_9)

# Check CHM crs
crs(TEAK_chm)

```

### Plot tower footprint and CHM.
```{r}

# Make raster a data frame
TEAK_chm_df <- as.data.frame(TEAK_chm, xy=TRUE)

# Plot the tower footprint shapefile onto the CHM
ggplot() +
  geom_raster(data = TEAK_chm_df,
              aes(x = x, y = y, fill = layer)) + 
  geom_sf(data = ECtower_TEAK_utm, color = "white", fill = NA) + 
  scale_fill_gradientn(name = "Canopy Height", colors = terrain.colors(10)) + 
  coord_sf(datum=utm11nCRS)

```

```{r, TEAK CHM summary: mean & sd}

# Get CHM data values (mean & sd) inside EC footprint shapefile
TEAK_fpCHM_mean <- raster::extract(x = TEAK_chm, y = ECtower_TEAK_utm, df=TRUE, fun=mean)
TEAK_fpCHM_mean
## mean = 10.111631

TEAK_fpCHM_sd <- raster::extract(x = TEAK_chm, y = ECtower_TEAK_utm, df=TRUE, fun=sd)
TEAK_fpCHM_sd
## sd = 10.97886

```  

```{r, summary eddy covariance}

max(flux$TEAK$data.fluxCo2.nsae.flux, na.rm=TRUE)
# 1412.851
min(flux$TEAK$data.fluxCo2.nsae.flux, na.rm=TRUE)
# -1028.751
mean(flux$TEAK$data.fluxCo2.nsae.flux, na.rm=TRUE)
# -0.5793503

```
